---
title: "Open Policing Project"
author: "Huyen Nguyen"
date: "3/27/2019"
output: html_document
---
##Setup 
Install the correct packages and load the libraries in.
If you have not installed the `tidyverse` package, make sure you uncomment the below lines.

```{r setup, message=F, warning=F}
#install.packages("tidyverse") 
#install.packages("lubridate")
library(tidyverse)
library(lubridate)

#load the data
hn <- read_csv("~/Desktop/Code/pa_philadelphia_2019_02_25.csv")
hn <- read_csv("https://datajournalism.tech/pa_philadelphia_2019_02_25.csv")
```

## Data Analysis
Explore the dataset provided by Stanford University. See more on their website
 <https://openpolicing.stanford.edu>.
 
```{r discover-the-dataset}
View(hn) #to view the data table
str(hn)  #to see the characteristics of variables
glimpse(hn) #to see a short summary of values in each column
colnames(hn) #to view column headers
hn <- hn %>% mutate(subject_sex=as.factor(subject_sex)) #convert chr to factor levels for the variable `subject_sex`
hn <- hn %>% mutate(subject_race=as.factor(subject_race)) #convert chr to factor levels for the variable `subject_race`
```
After viewing the dataset, you can analyze it to see the min, max, mean, median and other values for each variable. These are called descriptive statistics.

```{r run-basic-descriptive-statistics}
summary(hn)
```

There are some verbs that you need to memorize. See more at <https://learn.r-journalism.com/en/wrangling/dplyr/dplyr/>
First, the `select` verb helps you grab column(s) in a dataset

```{r select-one-or-more-columns}
a <- select(hn,subject_race, subject_age, subject_sex, type, contraband_found, arrest_made) #to select the subject_race columns and assign it to object a.
a <- hn %>% select(subject_race) #a different way to do the selection using %>% 

b <- filter(a, subject_age==10)

b <- group_by(b,arrest_made) %>% summarize(value=n())

```

Second, the `group_by` verb helps you categorize your values into fewer groups. The `summarize` verb always goes along with the `group_by` to help count the number of values for each group and compute the percentage of each group over the whole population.

```{r calculate-statistics-for-each-level-of-a-column }
sex <- hn %>% group_by(subject_sex) %>% 
  summarize(value=n(), prop = value/nrow(.))
# to count the numbers of stopped drivers for each gender and compute the percentage of column and assign it to a table called `sex`
View(sex) # to view the `sex` table

race <- hn %>% group_by(subject_race) %>% 
  summarize(value=n(), prop=value/nrow(.)) %>% 
  arrange(-value) # to count the numbers of stopped drivers for each race and compute the percentage of column and assign it to a table called `race`. Also, arrange the value in the descending order
View(race) # to view the `race` table

#Now it's your turn. Tell me how many stopped drivers were found with contraband? Or how many searches ended successfully? Write your code in the next line, without the hashtag.




```

##Data Visualization
We will need certain packages to be installed and called before creating our charts.

```{r data-viz-packages}
library(ggplot2)
#install.packages("devtools")
#devtools::install_github('bbc/bbplot')
library(bbplot)
```

###Line Chart
To create this chart, we will use the `ggplot2` package.
Let's create a simple line chart by modifying the code on the Stanford University's Open Policing Project website. This chart displays the total numbers of drivers who got stopped by the police over the year.

```{r line-chart}
line <- hn %>% 
  count(year = year(date), subject_race) %>% 
  ggplot(aes(x = year, y = n, color = subject_race)) +
  geom_point() +
  geom_line() +
  bbc_style()

line
```


###Bar Chart
```{r bar-chart}
#data prep
barprep <- hn %>% group_by(subject_race) %>% 
  summarize(value=n())

#make the plot

bar <- ggplot(barprep,
       aes(x=reorder(subject_race,value), y=value))+
  geom_bar(stat="identity", 
           position="identity", 
           fill="red")+
  geom_hline(yintercept = 0) +
  bbc_style()+
  labs(title="Stopped Drivers by Race",
       subtitle = "African American drivers got stopped the most in the city of Philadelphia,Pensylvania")+
  coord_flip()

options(scipen=10000)

bar
```


You can export the graphics by running the names of the objects in the Console and click the Export button under the Plots tab. Otherwise, you can use one of the following codes to export it:
```{r export-the-plot}
ggsave("bar.png", width=40, height=20, units= "cm")
ggsave("bar.svg", width=40, height=20, units= "cm") 
#ggsave only saves the last plot you created, so you may want to go up to the line chart and write ggsave("line.png")
```

###Pie Chart
This will be made with the ggplot2 package. 

```{r make-a-pie}
ggplot(barprep, aes(x="", y=value, fill=subject_race))+
  geom_bar(width=1, stat="identity")+
  coord_polar("y", start=0)
```

The following pie will be made with the `plotly` package.

```{r make-the-second-pie}
#install.packages("plotly")
library(plotly)

pieprep <- hn %>% group_by(subject_race) %>% 
  summarize(value=n()) %>% 
  arrange(-value)

color <- c("gray", "pink", "red", "yellow", "green")

pie <- plot_ly(data=pieprep, labels=~subject_race, values=~value, type="pie", textposition="inside",insidetextfont=list(color="white"),hoverinfo="text", text=~paste(subject_race), marker=list(colors=color, line=list(color="white", width=1)), showlegend=F)  %>% layout(title="Stopped Driver by Race")

pie

```

Now we will make a donut. How to? 

```{r make-a-donut}

plot_ly(data=pieprep, labels=~subject_race, values=~value, showlegend=F) %>% 
  add_pie(hole=.5) %>% 
  layout(title="Stopped Drivers by Race")

#The second way of creating a donut
donut <- hn %>% group_by(subject_race) %>% 
  summarise(value=n()) %>% 
plot_ly(labels=~subject_race, values=~value, showlegend=F) %>% 
  add_pie(hole=.5) %>% 
  layout(title="Stopped Drivers by Race")

donut

```


###Map with ggplot2
```{r map-w-ggplot2}
#list of packages we need
#install.packages("ggplot2")
#install.packages("ggmap")
#install.packages("maps")
#install.packages("mapdata")

#Call out the packages
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)

#get data for the base map
usa <- map_data("usa")
states <- map_data("state")
counties <- map_data("county")
world <- map_data("world2Hires")

#filter the year 2017
f <- filter(hn, year(date)==2017)

#filter Philadelphia, PA
philly <- filter(counties, region=="pennsylvania", subregion=="philadelphia")

#create the map

ggplot(philly) + geom_polygon(aes(x=long, y=lat, group=group))+
  coord_fixed(1.3)+
  geom_point(data=f, aes(x=lng, y=lat, color=subject_race), size=1)+ facet_wrap(~subject_race)

```
```{r}
#install.packages("httpuv")
#install.packages("leaflet")
library(httpuv)
library(leaflet)
m <- leaflet() %>% 
  addTiles() %>% 
  setView(lng= -75.172347, lat= 39.952150, zoom=16) %>% 
  addMarkers(lng= -75.172347, lat= 39.952150, popup="Philadelphia, PA")

m
```


```{r}
race <- colorFactor(c("pink", "black", "yellow", "red", "blue"), domain=c("white", "black", "asian/pacific islander", "hispanic", "other/unknown"), ordered=TRUE)
m2<- leaflet(f) %>%
  addProviderTiles(providers$CartoDB) %>% 
  setView(lng= -75.172347, lat= 39.952150, zoom=16) %>% 
  addCircleMarkers(~lng, ~lat, popup=paste("This is a", f$subject_race, "and", f$subject_sex, "driver."), weight= 3, radius=4, color=~race(subject_race), stroke=F, fillOpacity=.5)

m2
```




